---
title: "AIRLINE PASSENGER SATISFACTION"
author: "TEAM 7- Richik Ghosh, Shreya Sahay, Sanika Narayanapethkar, Vamsidhar Boddu "
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r}
library(ezids)
library(ggplot2)
library(gridExtra)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```

# Importing the Data

```{r,echo=TRUE,results='markup'}

data_pre<-data.frame(read.csv("/Users/richikghosh/Documents/Project Dats 6101/Airlines_train.csv"))
data<-data_pre
```

We have read the dataset using `read.csv()` and stored it in a dataframe called `data`.

# Structure of the data

```{r,echo=TRUE,results='markup'}
str(data)
```

We used `str()` function to study the structure of the data and further classified the variables into qualitative and quantitative types.

from the help menu, we can see that the variables are:
* `id`: customer id

* `Gender`: male or female

* `Customer.Type`: loyal or disloyal
* `Age`: tells us the age in years 
* `Type.Of.Travel`: personal or business
* `Class`: business, eco, or eco plus
* `Flight.Distance`: distance in nautical miles 
* `Inflight.wifi.service`: rating of wifi service
* `Departure.Arrival.time.convenient`: rating of time convenience 
* `Ease.of.Online.booking`: rating of ease of convenience 
* `Gate.location`: rating of gate location 
* `Food.and.Drink`: rating of food and drink items served in-flight
* `Online.boarding`: rating of boarding convenience
* `Seat.comfort`: rating of comfort 
* `Inflight.entertainment`: rating of how well customers were entertained 
* `On.board.service`: rating of on-boarding service 
* `Leg.room.service`: rating of space availability for legs
* `Baggage.handling`: rating of handling baggage
* `Checkin.service`: rating of check-in experience 
* `Inflight.service`: rating of in-flight experience 
* `cleanliness`: rating of cleanliness
* `Departure.Delay.in.Minutes`: flight departure delay in minutes
* `Arrival.Delay.in.minutes`: flight arrival delay in minutes  
* `satisfaction`: rating of overall experience 

**NOTE: All the ratings are on a scale of 0-5**


These variables can be re-classified as:

* Qualitative: `satisfaction`, `Class`, `Type.of.Travel`, `Gender`, `Customer.Type`

* Quantitative: `id`, `Age`, `Flight.Distance`, `Inflight.wifi.service`, `Departure.Arrival.time.convenient`, `Ease.of.Online.booking`, `Gate.location`, `Food.and.Drink`, `Online.boarding`, `Seat.comfort`, `Inflight.entertainment`, `On.board.service`, `Leg.room.service`, `Baggage.handling`, `Checkin.service`, `Inflight.service`, `cleanliness`, `Departure.Delay.in.Minutes`, `Arrival.Delay.in.minutes`.

# Data Preprocessing

## Dropping columns

```{r,echo=TRUE}
data<- subset(data_pre, select = -c(X, id))
```

Using the `subset()` function we removed the `id` and `X` columns since they were helpful only for data retrieval/identification and do not contribute much to data analysis.

## Factoring the variables
```{r,echo=TRUE,results='markup'}
data$satisfaction<-as.factor(data$satisfaction)
data$Class<-as.factor(data$Class)
data$Type.of.Travel<-as.factor(data$Type.of.Travel)
data$Gender<-as.factor(data$Gender)
data$Customer.Type<-as.factor(data$Customer.Type)
```

We did factoring of the qualitative variables for performing EDA. We converted `satisfaction`, `Class`, `Type.of.Travel`, `Gender` and `Customer.Type`. We use the `as.factor()` function.


## Replacing the NA values with mean

```{r,echo=TRUE,results='markup'}
data$Arrival.Delay.in.Minutes[is.na(data$Arrival.Delay.in.Minutes)] = mean(data$Arrival.Delay.in.Minute,na.rm = TRUE)
paste("The NA's in the dataset after Arrival Delay in Minute with mean is:",sum(is.na(data)))
```
 

We replaced the NA values with the mean. We used `sum(is.na())` method to calculate the `mean`. 

## Removal of Outliers
```{r,echo=TRUE,results='markup'}
data2<-outlierKD2(data,Flight.Distance,rm=TRUE,boxplt = TRUE,qqplt = TRUE)
data3<-outlierKD2(data,Departure.Delay.in.Minutes,rm=TRUE,boxplt = TRUE,qqplt = TRUE)
data4<-outlierKD2(data3,Arrival.Delay.in.Minutes,rm=TRUE,boxplt = TRUE,qqplt = TRUE)
```

## Categorization of Age Groups

```{r,echo=TRUE}
data <- within(data,{Age.cat = NA
  Age.cat[Age<=20] = "Under 20"
  Age.cat[Age>=21 & Age<= 40 ] = "20-40"
  Age.cat[Age>=41 & Age<= 80 ] = "40-80"
  Age.cat[Age>=80 ] = "above 80"})
data$Age.cat<- factor(data$Age.cat,levels= c("Under 20","20-40","40-80","above 80"))
```
```{r}
data <- within(data,{Age.cat2 = NA
  Age.cat2[Age<=20] = 1
  Age.cat2[Age>=21 & Age<= 40 ] = 2
  Age.cat2[Age>=41 & Age<= 80 ] = 3
  Age.cat2[Age>=80 ] = 4})
#data$Age.cat2<- factor(data$Age.cat2,levels= c("1","2","3","4"))
```

We created a new categorical variable named `Age.cat` based on the values of the existing `Age` variable. It categorizes individuals into age groups and then converts this new variable into a factor variable with specified levels. `within()` function is used to form new variable within the existing `dataset`.

# Summary of the data

```{r,echo=TRUE,results='markup'}
paste("The NA's in the dataset column Arrival.Delay.in.Minutes is:",sum(is.na(data$Arrival.Delay.in.Minutes)))
summary(data)
```



These lines of code generate a message that combines a text string with the `count` of `NA values` using the `is.na()` function in the `Arrival.Delay.in.Minutes` variable. The `summary()` function provides a statistical summary of the entire dataset.



# Visualizations


## Age distribution
```{r,echo=TRUE,results='markup'}

ggplot(data, aes(x=Age)) +  geom_density(fill="#9FA8DA", color="#9FA8DA", alpha=0.5)+ theme_bw() + theme()+ ggtitle("Passenger Age Distribution")+labs(title = "Passenger Age Distribution",
       x = "Age",
       y = "Density of the Customers")
```
## Barplot of Gender vs Satisfaction

```{r,echo=TRUE,results='markup'}



## Gender vs Satisfaction plot using bar plot

gender_1 <- table(data$Gender)
gender_dist <- as.data.frame(gender_1)
names(gender_dist)<-c("Gender","Count")
                      
ggplot(gender_dist, aes(x = Gender, y = Count, fill = Gender)) +
  geom_bar(stat = "identity") +
  labs(title = "Gender Distribution",
       x = "Gender",
       y = "Number of Customers") +
  scale_fill_manual(values = c("Male" = "#F06292", "Female" = "#9FA8DA", "Other" = "gray")) +
  theme_minimal()
```

## Barplot of Satisfaction

```{r, echo=TRUE,results="show"}
## Overall satisfaction insight of the passengers

ggplot(data, aes(x=satisfaction,fill=satisfaction))+geom_bar()+ggtitle("Total Satisfied and Neutral or Unsatisfied customers in the Dataset")+ theme_bw() +
  theme()+ ylab("Total Customers") + xlab("Feeling of the customer")+ theme(
    axis.title.x = element_text(color = "#9FA8DA"),  
    axis.title.y = element_text(color = "#F06292"),  
    plot.title = element_text(color = "#AD1457"),   
    axis.text.x = element_text(color = "black"))+ scale_fill_manual(values = c("satisfied" = "#F06292", "neutral or dissatisfied" = "#9FA8DA", "Other" = "gray"))
```


## Barplot of dependencies of three varibles on Satisfaction of a customer

```{r, echo=TRUE,results="show"}

## Type of Travel vs Satisfaction plot using bar plot
ggplot(data, aes(x=Type.of.Travel,fill=satisfaction))+geom_bar(position = "dodge")+ggtitle("Customer Satisfaction according to Type of Travel")+ theme_bw() +
  theme()+ylab("Customer Volume") + xlab("Type of travel")+ theme(
    axis.title.x = element_text(color = "black"),  
    axis.title.y = element_text(color = "black"),  
    plot.title = element_text(color = "black"),   
    axis.text.x = element_text(color = "black"))+ scale_fill_manual(values = c("satisfied" = "#F06292", "neutral or dissatisfied" = "#9FA8DA", "Other" = "gray"))

ggplot(data, aes(x=Class,fill=satisfaction))+geom_bar(position = "dodge")+ggtitle("Customer Satisfaction based on Class")+ theme_bw() +
  theme()+ylab("Number of Cusotmers") + xlab("Class")+ theme(
    axis.title.x = element_text(color = "black"),  
    axis.title.y = element_text(color = "black"),  
    plot.title = element_text(color = "black"),   
    axis.text.x = element_text(color = "black"))+ scale_fill_manual(values = c("satisfied" = "#F06292", "neutral or dissatisfied" = "#9FA8DA", "Other" = "gray"))

ggplot(data, aes(x=Customer.Type,fill=satisfaction))+geom_bar(position = "dodge")+ggtitle("Customer Satisfaction according to Customer Type")+ylab("Number of Cusotmers") + xlab("Customer type")+theme_bw()+theme(
    axis.title.x = element_text(color = "black"),  
    axis.title.y = element_text(color = "black"),  
    plot.title = element_text(color = "black"),   
    axis.text.x = element_text(color = "black"))+ scale_fill_manual(values = c("satisfied" = "#F06292", "neutral or dissatisfied" = "#9FA8DA", "Other" = "gray"))

#grid.arrange(p1, p2, p3, p4, p5)

```

We called the packages `ggplot2` and `gridExtra` using the `library()` function.

We count the frequency of each unique value in the `Gender` column and store this information in the gender_1 table.
We use the `as.data.frame()` to convert the tabulated data into a format that's more suitable for data analysis.
`names(gender_dist) <- c("Gender", "Count")` this step is often done to make the data frame more descriptive and easy to work with.

We created a `barplot` using `ggplot()` that visualizes the `Gender` distribution based on the data in the gender_dist data frame. `stat =identity` argument indicates that the heights of the bars should be taken directly from the `Count` column in the data frame.

We created a `density plot` using `ggplot()` that visualizes the distribution of `Age` in the `data` data frame.  The `alpha` argument controls the transparency of the density curve which set to .

We created a `barplot` using `ggplot()` that visualizes the `satisfaction` (different levels of satisfaction) distribution based on the data in the `gender_dist` data frame.

We created a `barplot` using `ggplot()` that compares `satisfaction` with different `Types.of.Travel` from the `data` data frame.

We created a `barplot` using `ggplot()` that compares `satisfaction` for various `Classes` in the `data` dataset.

We created a `barplot` using `ggplot()` that compares `satisfaction` for various `Customer.Type` in the `data` dataset.


## Barplot of Customer Satisfaction based on different Age Groups
```{r,echo=TRUE,results='markup'}
ggplot(data, aes(x=Age.cat,fill=satisfaction))+geom_bar(position = "dodge")+ggtitle("Customer Satisfaction based on different Age Groups")+ylab("Number of Customers") + xlab("Age Category")+theme_bw()+theme(
    axis.title.x = element_text(color = "black"),  
    axis.title.y = element_text(color = "black"),  
    plot.title = element_text(color = "black"),   
    axis.text.x = element_text(color = "black"))+ scale_fill_manual(values = c("satisfied" = "#F06292", "neutral or dissatisfied" = "#9FA8DA", "Other" = "gray"))
```

We created a `barplot` using `ggplot()` that visualizes the `satisfaction` distribution based on the data in the `Age.cat` data frame.

It can be deduced from the above bar plot that the age group that is under 20 and above 60 are compartively neutral or dissatisfied with the flight experience.

## Boxplot age vs satisfaction of a customer
```{r,echo=TRUE,results='markup'}
ggplot(data = data, aes(x=as.character(satisfaction ), y=Age, fill=satisfaction)) +
    geom_boxplot() +
    labs(title="Boxplot of age and Customer Satisfaction", x="Customer Satisfaction", y="Age")+ scale_fill_manual(values = c("satisfied" = "#F06292", "neutral or dissatisfied" = "#9FA8DA", "Other" = "gray"))+ scale_y_continuous(limits = c(0, 80))
```


```{r,echo=TRUE}
data_sat=subset(data4, data$satisfaction=='satisfied')
data_dissat=subset(data4, data$satisfaction=='neutral or dissatisfied')
```

## Boxplot of departure delay vs satisfaction

```{r,echo=TRUE,results='markup'}
ggplot(data4, aes(y = Departure.Delay.in.Minutes, fill = satisfaction)) +
  geom_boxplot() +  
  scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3")) + 
  labs(title = "Boxplot of departure delay vs satisfaction",
       x = "Satisfaction", y = "Departure Delay (in Minutes)") +
  theme_minimal()+ scale_y_continuous(limits = c(0, 30))
```
## Boxplot of arrival delay vs satisfaction

```{r,echo=TRUE,results='markup'}
ggplot(data4, aes(y = Arrival.Delay.in.Minutes, fill = satisfaction)) +
  geom_boxplot() +  
  scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3")) + 
  labs(title = "Boxplot of arrival delay vs satisfaction",
       x = "Satisfaction", y = "Arrival delay (in Minutes)") +
  theme_minimal()+ scale_y_continuous(limits = c(0, 30))
```

We created a `boxplot` using `ggplot()` that compares `satisfaction` with `Age`.


## VIOLIN CHART AND CORREALATION COEFFICIENT

```{r,echo=TRUE,results='markup'}
library(ggplot2)

ggplot(data, aes(x = satisfaction, y = Inflight.entertainment, fill = satisfaction)) +
  geom_violin() +
  labs(title = "Overall Satisfaction vs. In-flight Entertainment Quality",
       x = "Overall Satisfaction",
       y = "In-flight Entertainment Quality")+scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))


ggplot(data, aes(x = satisfaction, y = Cleanliness, fill = satisfaction)) +
  geom_violin() +
  labs(title = "Overall Satisfaction vs. Cleanliness",
       x = "Overall Satisfaction",
       y = "Cleanliness of the flight")+scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))


ggplot(data, aes(x = satisfaction, y = Food.and.drink, fill = satisfaction)) +
  geom_violin() +
  labs(title = "Overall Satisfaction vs. Food and Drinks",
       x = "Overall Satisfaction",
       y = "Food and Drinks")+scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))


ggplot(data, aes(x = satisfaction, y = Checkin.service, fill = satisfaction)) +
  geom_violin() +
  labs(title = "Overall Satisfaction vs. Check in Service",
       x = "Overall Satisfaction",
       y = "Check in Service")+scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))



ggplot(data, aes(x = satisfaction, y = Seat.comfort, fill = satisfaction)) +
  geom_violin() +
  labs(title = "Overall Satisfaction vs. Seat Comfort",
       x = "Overall Satisfaction",
       y = "Seat comfort")+scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))



cor_Inflight <- cor(data$Inflight.entertainment, as.numeric(factor(data$satisfaction)))
print(paste("Correlation coefficient based on Overall Satisfaction and Inflight Entertainment:", cor_Inflight))

cor_clean <- cor(data$Cleanliness, as.numeric(factor(data$satisfaction)))
print(paste("Correlation coefficient based on Overall Satisfaction and Cleanlimess:", cor_clean))

cor_food <- cor(data$Food.and.drink, as.numeric(factor(data$satisfaction)))
print(paste("Correlation coefficient based on Overall Satisfaction and Food and Drinks:", cor_food))

cor_checkin <- cor(data$Checkin.service, as.numeric(factor(data$satisfaction)))
print(paste("Correlation coefficient based on Overall Satisfaction and Check in Service:", cor_checkin))

cor_seat <- cor(data$Seat.comfort, as.numeric(factor(data$satisfaction)))
print(paste("Correlation coefficient based on Overall Satisfaction and Seat Comfort:", cor_seat))
```

We created `violin plots` using `ggplot()` and calculate the correlation coefficients between
`Satisfaction` and different aspects of the flight experience.


We use the `geom_violin()` function in `ggplot()` to create a plot that shows the distribution of `Inflight.Entertainment` for each level of `Satisfaction`. Then we print the correlation coefficient between `Inflight.entertainment` and `satisfaction`using the `cor()` function.

We use the `geom_violin()` function in `ggplot()` to create a plot that shows the distribution of `Cleanliness` for each level of `Satisfaction`. Then we print the correlation coefficient between `Cleanliness` and `satisfaction`using the `cor()` function.

We use the `geom_violin()` function in `ggplot()` to create a plot that shows the distribution of `Food.and.drink` for each level of `Satisfaction`. Then we print the correlation coefficient between `Food.and.drink` and `satisfaction`using the `cor()` function.


We use the `geom_violin()` function in `ggplot()` to create a plot that shows the distribution of `Checkin.service` for each level of `Satisfaction`. Then we print the correlation coefficient between `Checkin.service` and `satisfaction`using the `cor()` function.

We use the `geom_violin()` function in `ggplot()` to create a plot that shows the distribution of `Seat.comfort` for each level of `Satisfaction`. Then we print the correlation coefficient between `Seat.comfort` and `satisfaction`using the `cor()` function.


## HISTOGRAM of flight distance travelled by a customer
```{r,echo=TRUE,results='markup'}
ggplot(data2,aes(x=Flight.Distance,fill=satisfaction))+geom_histogram(binwidth=500)+labs(title="Histogram of frequency of Flight Distance based on satisfaction",x="Flight.Distance",y="Frequency")+theme_minimal()+ theme(
    axis.title.x = element_text(color = "black"),  
    axis.title.y = element_text(color = "black"),  
    plot.title = element_text(color = "black"),   
    axis.text.x = element_text(color = "black"))+scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))
```

We will make a `histogram` using `ggplot()` to visualize the relationship between `satisfaction` and `Flight.Distance`.



## BOXPLOT of flight distance and customer satisfaction
```{r,echo=TRUE,results='markup'}
ggplot(data = data2, aes(x=as.character(satisfaction ), y=Flight.Distance, fill=satisfaction)) +
    geom_boxplot() +
    labs(title="Boxplot of Flight Distance vs Customer Satisfation ", x="Customer Satisfation ", y="Flight Distance")+scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))
```
`boxplot` using `ggplot()` to visualize the relationship between `satisfaction` and `Flight.Distance`.

## Box Plot of Flight Distance by Class and Satisfaction Level
```{r,echo=TRUE,results='markup'}

ggplot(data2, aes(x = Class, y = Flight.Distance, fill = satisfaction)) +
  geom_boxplot() +
  labs(x = "Class", y = "Flight Distance", title = "Box Plot of Flight Distance by Class and Satisfaction Level") + scale_fill_manual(values = c("satisfied" = "lightblue3", "neutral or dissatisfied" = "lightpink3"))

```

## Correlation Plot of the dataset

```{r,echo=TRUE,results='markup'}
library(corrplot)

data$satisfaction_numeric <- ifelse(data$satisfaction == "satisfied", 1, 0)

var_columns <- c("Flight.Distance", "Inflight.wifi.service", "Departure.Arrival.time.convenient",
                      "Ease.of.Online.booking", "Gate.location","Age", "Food.and.drink", "Online.boarding",
                      "Seat.comfort", "Inflight.entertainment", "On.board.service", "Leg.room.service",
                      "Baggage.handling", "Checkin.service", "Inflight.service", "Cleanliness","Arrival.Delay.in.Minutes","Departure.Delay.in.Minutes","satisfaction_numeric")

correlation_matrix <- cor(data[var_columns])

corrplot.mixed(correlation_matrix, tl.col = "black", tl.srt = 45,
               tl.cex = 0.9, tl.offset = 1, tl.pos = "lt", lower = 'number', upper = 'pie', order = 'hclust')

```


# Statistical Tests

## CHI-SQUARE TEST ON THREE VARIABLES
```{r,echo=TRUE,results='markup'}
alpha=0.05
ctest1 = chisq.test(data$satisfaction, data$Type.of.Travel)

ctest2 = chisq.test(data$satisfaction, data$Class)

ctest3 = chisq.test(data$satisfaction, data$Customer.Type)

ctest1
ctest2
ctest3
```



We have conducted `chisq.test()` to find the $\chi^2$ analysis on three variables - `Type.of.Travel`, `Class`, `Customer.Type`.

Analysis is as follows:


Null Hypothesis ($H_{0}$):
There is no significant association between `Satisfaction` and `Type.of.Travel`, `Class`, `Customer.Type`, respectively.

Alternative Hypothesis($H_{A}$):
There is a significant association between `Satisfaction` and `Type.of.Travel`, `Class`, `Customer.Type`, respectively.


The `p-value` returned by the $\chi^2$ test in comparison of satisfaction and `Type.of.Travel` is  $`r ctest1$p.value`$ then satisfaction and `Class` is  $`r ctest2$p.value`$ and finally satisfaction and `Customer.Type` is  $`r ctest3$p.value`$, which are less than the standard  $\alpha$=0.05. As a result, the null hypothesis is rejected, and the alternate hypothesis is accepted. We may draw the conclusion that there is significant association between `Satisfaction` and `Type.of.Travel`,  `Class`, `Customer.Type`, respectively.


## T-test on `age` variable

```{r,echo=TRUE,results='markup'}
a<-t.test(Age~satisfaction,data = data)
a
```

We performed a T-Test using the `t.test()` function and it can be observed that the means of the neutral/dissatisfied and satisfied are different with respective values of 37.6 and 41.8. The `p-value` of $`r a$p.value`$ is less than $\alpha$=0.05 which indicates that the dissatisfied/neutral and satisfied customers have different age groups

## Anova test on age category and satisfaction

```{r,echo=TRUE,results='markup'}
model<-aov(Age.cat2 ~ satisfaction,data=data)
ezids::xkabledply(model,title="Anova test result of Age and Satisfaction")

```
We performed an ANOVA test on satisfaction and age category using `aov()` function. We found that the p value is $`r summary(model)[[1]][["Pr(>F)"]][1]`$ is lower than $\alpha$=0.05, which means we can reject the null hypothesis of that satisfaction of customer staying the same around each age category ("Under 20", "20-40", "40-80", and "above 80"). The means of satisfaction scores for each of these age groups are compared using the ANOVA test to see if there are any statistically significant differences.


## T test on arrival and departure delay

```{r,echo=TRUE,results='markup'}
b<-t.test(data4$Arrival.Delay.in.Minutes,data4$Departure.Delay.in.Minutes)
print(b)
```

Analysis is as follows:
Null Hypothesis($H_{0}$):
The mean of departure and arrival delay time are same 

Alternative Hypothesis($H_{A}$):
The mean of departure and arrival delay time are different 

The `p-value` returned by comparing arrival delay and departure delay is $`r b$p.value`$ which is lower than $\alpha$=0.05. Thus, we can reject the null hypothesis and say departure and arrival delay time are different.

## T-test on arrival and departure delay and satifaction


```{r,echo=TRUE,results='markup'}
b1<-t.test(Arrival.Delay.in.Minutes~satisfaction,data = data4)
b2<-t.test(Departure.Delay.in.Minutes~satisfaction,data = data4)


print(b1)
print(b2)

```


We first conducted an independent t-test using the `t.test()` function to determine if the two 
variables `Arrival.Delay.in.Minutes` and `Departure.Delay.in.Minutes` have similar means.


Analysis is as follows:
Null Hypothesis($H_{0}$):
The mean of departure and arrival delay time is the same across all levels of satisfaction.

Alternative Hypothesis($H_{A}$):
The mean of departure and arrival delay time is different for at least one pair of satisfaction levels.

The `p-value` returned for arrival delay by the $t-test$ is $`r b1$p.value`$ and for departure delay by the $t-test$ is $`r b2$p.value`$. As a result, the null hypothesis is rejected, and the alternate hypothesis is accepted. We draw the conclusion that the mean is different for at least one pair of satisfaction levels.


## CHI-SQUARE TEST on satisfaction and inflight details
```{r,echo=TRUE,results='markup'}
ctest1 = chisq.test(data$satisfaction, data$Inflight.entertainment)
ctest2 = chisq.test(data$satisfaction, data$Cleanliness)
ctest3 = chisq.test(data$satisfaction, data$Food.and.drink)
ctest4 = chisq.test(data$satisfaction, data$Seat.comfort)

print(ctest1)
print(ctest2)
print(ctest3)
print(ctest4)
```



We have conducted `chisq.test` to find the $\chi^2$ on four variables - `Inflight.entertainment`, `Cleanliness`, `Food.and.drink`, `Seat.comfort`.

Analysis is as follows:


Null Hypothesis($H_{0}$):
There is no significant association between `Satisfaction` and `Inflight.entertainment`, `Cleanliness`, `Food.and.drink`, `Seat.comfort` respectively.

Alternative Hypothesis($H_{A}$):
There is a significant association between `Satisfaction` and `Inflight.entertainment`,  `Cleanliness`, `Food.and.drink`, `Seat.comfort` respectively.


The `p-value` returned by the $\chi^2$ test on `Inflight.entertainment`,  `Cleanliness`, `Food.and.drink`, `Seat.comfort` is `p-value` of $`r ctest1$p.value`$, $`r ctest2$p.value`$, $`r ctest3$p.value`$, $`r ctest4$p.value`$ which are lesser than $\alpha$=0.05 respectively. As a result, the null hypothesis is rejected, and the alternate hypothesis is accepted. We may draw the conclusion that there is significant association between `Satisfaction` and `Inflight.entertainment`,  `Cleanliness`, `Food.and.drink`, `Seat.comfort` respectively.


## T test on flight distance and satisfaction
```{r,echo=TRUE,results='markup'}
k<-t.test(Flight.Distance~satisfaction,data = data)
```



We first conducted an independent t-test using the `t.test()` function to determine if the two 
variables `Flight.Distance` and `satisfcation` have similar means.
Analysis is as follows:

Null Hypothesis($H_{0}$):
The mean flight distance is the same across all levels of satisfaction


Alternative Hypothesis($H_{A}$):
The mean flight distance is different for at least one pair of satisfaction levels.


The `p-value` returned by the $t-test$ is `p-value` of $`r k$p.value`$ is less than $\alpha$=0.05. As a result, the null hypothesis is rejected, and the alternate hypothesis is accepted. We may draw the conclusion that the mean is different for at least one pair of satisfaction levels.

